---
title: "Report"
author: "Yue Li, Suofei Feng"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(fda.drug.stats290)
```

# Introduction

The FDA offers a trove of valuable data through their Open FDA website. However, 
the API can often be tedious and difficult to navigate. This package provides 
researchers with a fast and easy way to grab the API data and avoid overhead. More 
information about the Open FDA initiative can be found here, on their [site](open.fda.gov).
This package provides users with an easy option to gather data using parametrized queries
and reshape the data for easier use. We also provide functions for plotting/graphing 
and gathering summary values. We came across a few road blocks while working on this
project. There were a 

This document introduces you to fda.drug.stats290's basic  set of tools and shows 
you how to apply them in different ways.

# Use cases

A user may want to explore a subset of serious reactions to drugs. First, we can
try to understand the summary of what might be going on. Perhaps we can first understand
when serious reactions to drugs would occur, and what kinds of drugs they might be.

```{r}
count_events("receivedate")
visualize_events("receivedate")

visualize_events("patient.drug.openfda.generic_name")
```

Once we understand the brief summary of what is going on, we can dig into the
data.

```{r}
t_1 <- fda_api_table(serious = 1, limit = 50)
```

Using `t_1`, we can do a few things:

```{r}
t_1_df <- fda_dataframe(t_1) #get a data frame from t_1 without openfda and drug data
head(t_1_df)
```

```{r}
t_1_drug <- fda_drug_dataframe(t_1) #get a data frame from t_1 for drug without openfda data
head(t_1_drug)
```

```{r}
openfda_t_1 <- get_openfda(t_1)
summary(openfda_t_1$`1`) #show the openfda data of the patient with ID = 1
```

```{r}
manufact_count <- openfda_count("manufacturer_name", openfda_t_1)
head(manufact_count)
```
Visualize the top reactions of the top 3 most frequent activesubstance names:
```{r}
top_react <- view_top_react(t_1_df, t_1_drug, top_name = 3, top_react = 8)
head(top_react)
```


# Contributions

Yue contributed to the functions visualize_events, count_events, and fda_api_table,
as well as gathering the dataset and creating the documentation for the report.

Suofei created the functions fda_dataframe, fda_drug_dataframe, get_openfda, 
openfda_count, and view_top_react.

# Additional information and ways to use the functions

In order to familiarize yourself with the dataset, please read the 
[variable documentation](https://open.fda.gov/apis/drug/event/searchable-fields). 
This will be useful for understanding the values of the many variables that FDA provides, 
and will be necessary for providing search filters and count queries.

## Data: drug_events

This is a static dataset pulled from Q1 of 2018 (part 1), which contains over 
12,000 instances of adverse drug events. This is a way for researchers to play 
with and understand the basic structure of the FDA dataset. This dataset will
not be updated, so in general, we recommend using the fda_api_table function
to obtain your dataset. 

## fda_api_table

This is a function that allows people to return tables from the FDA API. The 
structure of this dataset is similar to the static dataset stored in drug_events.
It primarily provides a person with the ability to set a start and end date. 
Further, we provide assistance with declaring some filter variables. 


The function allows you to filter based on a few key features which are listed in
more detail on the help page. Because the FDA dataset is so vast and large, we
limited the search to the key features. In the future, we can consider expanding
the filter to include all of the variables. However, in this case, to provide
ease of use, we felt that restricting the filter to certain key words would be enough.

```{r}
t_1 <- fda_api_table()
  
t_2 <- 
  fda_api_table(
    limit = 100, 
    start_date = "2016-01-01", 
    end_date = "2018-01-01",
    serious = 1,
    generic_name  = "asprin"
  )
```


This package also provides the ability to pass in your API key. The API key
allows you to obtain more requests per day than just your IP address. To learn
about the API key, look at their [authentication page](https://open.fda.gov/apis/authentication/).
In this case, you are able to pass in your authentication key simply by passing
in your key in the api_key field.

```{r, eval=FALSE}
t <- count_events("receivedate", api_key = "YOUR API KEY HERE")
```


## count_events and visualize_events

One useful way to also look at the FDA data is to count the total value of certain
events using the function count_events.

```{r}
count_events("receivedate")

count_events(
  "receivedate", 
  limit = 100,
  start_date = "2016-01-01",
  end_date = "2018-01-01",
  serious = 1,
  generic_name  = "asprin"
)
```

For the function visualize_events, we build on top of the count_events function
and create visualizations from the FDA dataset. Because many of the result returns
thousands of values, which will be impossible to graph, we recommend users to limit
the number of values to the top 20 and set a hard limit at 50 values, unless the
variable is a continuous variable like time. 

```{r}
# visualize all patient weight events
visualize_events("patient.patientweight")
# visualize receivedates where patientsex was male and was submitted by a pharmacist
visualize_events("receivedate", patientsex = 1, qualification = 2)
# visualize by gender (0 = unknown, 1 = male, 2 = female)
visualize_events("patient.patientsex")
# visualize by top 30 patient reactions
visualize_events("patient.reaction.reactionmeddrapt.exact", limit = 30)
```


## fda_dataframe and fda_drug_dataframe

One challenge of the fda_api_table that is called is that many of the values are
deeply nested inside of one another. For example, to find openfda data, one 
needs to go from fda_api_table's surface, to patient level, and to drug level 
under patient. Then one can see openfda as data frames, whose entries are 
lists. The functions, fda_dataframe, and fda_drug_dataframe help flatten the 
tables returned by fda_api_table (or the in-built data set). The reason we 
separate the general data frame for patient information and the drug data 
frame for the patient is that, if we combine the two data frames, due to the 
large number of levels of the data nesting, there will be large quantities of 
data replication inside one data frame, which will make the data frame 
difficult to read.
```{r}
t_2_df <- fda_dataframe(t_2)
head(t_2_df)
```

```{r}
t_2_drug <- fda_drug_dataframe(t_2)
head(t_2_drug)
```

In case the user wants to combine the information from the two data frames, 
same objects returned by fda_api_table should be fed into the two functions.

##get_openfda and openfda_count
As the openfda data dwells as separate data frames of data frames at a deeper 
level than drug information, it is more efficient to extract them into a list 
with a function get_openfda:
```{r}
openfda_data <- get_openfda(t_2)
summary(openfda_data$`2`) #show the summary of openfda data of the patient with ID=2
openfda_data_small <- get_openfda(t_2, id = 2) #user can specify whose openfda data is needed using the argument id
summary(openfda_data_small)
```

An application of the output of get_openfda is to get the counts of a 
specified term by calling openfda_count. The output is a data frame. There are
two required arguments: var_name, which specifies the term (must be
meaningful, one of ("generic_name", "brand_name", "manufacturer_name", 
"substance_name", "product_type", "route")); openfda_data, which is returned 
by get_openfda.

```{r}
brand_count <- openfda_count("brand_name", openfda_data)
head(brand_count)
```

A possible application of the output of this openfda_count would be Word 
Cloud. For example, with proper parameters, the frequencies of each brand name
can be easily visualized by Word Cloud using the package wordcloud and seeing 
what the most frequently used term is.

```{r, eval = FALSE}
library(wordcloud)

wordcloud(words = brand_count$word, freq = brand_count$count, scale = c(1, .2), min.freq = 10, max.words=50,random.order=FALSE, random.color = FALSE, rot.per=0.35, colors=brewer.pal(8, "Dark2"), fixed.asp = TRUE)
```

##view_top_react
With the shared ID when producing fda_dataframe and fda_drug_dataframe using 
the same fda_api_table object, the user can combine the two data frames to get
some insights. view_top_react is an example which returns a data frame with 
activesubstance from fda_drug_dataframe object, and its corresponding patient
reaction from fda_dataframe object. It also automatically shows histograms of
the counts of reactions for each activesubstance. The number of 
activesubstance names for display, and the number of reactions to be shown 
explicitly (not in "others"), can be specified with arguments top_name 
(activesubstance, default = 5) and top_react (reaction, default = 8).

```{r}
subs_reaction <- view_top_react(t_2_df, t_2_drug, top_name = 6, top_react = 10)
```

```{r}
head(subs_reaction)
```

